`infercnv` results exploration for `r params$sample_id`

### 2024-11-04 analyses/cell-type-wilms-tumor-06/notebook_template/06_cnv_infercnv_exploration.Rmd

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warnings=FALSE
                      )

```

@@

```{r}

subdiagnosis <- readr::read_tsv(
  file.path("..", "..", "..", "data", "current", params$scpca_project_id, "single_cell_metadata.tsv"),
  show_col_types = FALSE
  ) |>
  dplyr::filter(scpca_sample_id == params$sample_id) |>
  dplyr::pull(subdiagnosis)

```

@@

```{r}

library(infercnv)
library(SCpubr)
library(ggplot2)
library(patchwork)
library(Seurat)
```

@@

```{r}

Do_CNV_heatmap <- function(object, infercnv_obj, group.by, reference_value){

  out <- SCpubr::do_CopyNumberVariantPlot(sample = object,
                                        infercnv_object = infercnv_obj,
                                        using_metacells = FALSE,
                                        chromosome_locations = SCpubr::human_chr_locations,
                                        return_object = FALSE,
                                        group.by = group.by
                                        )
  out <- out + 
    ggtitle(glue::glue("Copy Number Variant Plot, ", reference_value)) +
    ylab(label = "")

  return(out)
}

```

@@

```{r}

Do_CNV_score <- function(seurat_oject,infercnv_obj,reference_value){

  score <- apply(infercnv_obj@expr.data,2,function(x){ sum(x < 0.95 | x > 1.05)/length(x) })

@@

```{r }
visualize_metadata <- function(object, meta, group.by){

  if(is.numeric(object@meta.data[,meta])){
     d <- SCpubr::do_FeaturePlot(object, 
                                              features = meta, 
                                              pt.size = 0.2, 
                                              legend.width = 0.5, 
                                              legend.length = 5, 
                                              legend.position = "right") + ggtitle(meta)
    b <- SCpubr::do_ViolinPlot(object, 
                                             features = meta, 
                                             ncol = 1, 
                                             group.by = group.by, 
                                             legend.position = "none")

   return(d + b + plot_layout(ncol = 2, widths = c(2,4))) 
  }

  else{

  d <- SCpubr::do_DimPlot(object, reduction="umap", group.by = group.by, label = TRUE, repel = TRUE) + ggtitle(paste0(meta," - umap")) + theme(text=element_text(size=18))
  b <- SCpubr::do_BarPlot(sample = object,
                         group.by = meta,
                         split.by = group.by,
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(print(group.by)) +
                         theme(text=element_text(size=18))
  return(d + b + plot_layout(ncol = 2, widths = c(2,4)))
  }

}
```

@@

```{r }
visualize_feature <- function(object, feature, group.by ){

                  d <- SCpubr::do_FeaturePlot(object, 
                                              features = feature, 
                                              pt.size = 0.2, 
                                              legend.width = 0.5, 
                                              legend.length = 5, 
                                              legend.position = "right") + ggtitle(as.character(feature))
                  b <- SCpubr::do_ViolinPlot(object, 
                                             features = feature, 
                                             ncol = 1, 
                                             group.by = group.by, 
                                             legend.position = "none",
                                             assay = "SCT") + ylab(as.character(feature))

                  return(d + b + plot_layout(ncol = 2, widths = c(2,4))) 
}
```

@@

```{r }
visualize_density <- function(object, features, group.by ){

                  d <- SCpubr::do_RidgePlot(object, 
                                            feature = features, 
                                            group.by = group.by, 
                                             legend.position = "none",
                                             assay = "SCT") + ylab(features)

                  return(d) 
}

```

@@

```{r}

wrapper_explore_hmm <- function(infercnv_obj, cnv_threshold = 1){

  p <- list()

@@

  #### Feature plot and repartition of the CNV per chromosome
  for(i in 1:22){
    tmp1 <- visualize_feature(infercnv_obj, feature=glue::glue("proportion_cnv_chr", i), group.by = "fetal_kidney_predicted.compartment")
    tmp2 <- visualize_density(infercnv_obj, features=glue::glue("proportion_cnv_chr", i), group.by = "fetal_kidney_predicted.compartment")
    p[[glue::glue("proportion_cnv_chr", i)]] <- tmp1 + tmp2 + plot_layout(ncol = 3, widths = c(1,2,2))
  }

  #### DotPlot
  p[["Dotplot"]] <- DotPlot(infercnv_obj, features =  colnames(infercnv_obj@meta.data)[grepl("has_cnv_chr", colnames(infercnv_obj@meta.data))], group.by = "fetal_kidney_predicted.compartment", assay = "SCT") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  #### CNV score
  meta <- infercnv_obj@meta.data
  meta$has_cnv_score <- rowSums(meta[,grepl("has_cnv_chr", colnames(meta))])
  meta$has_cnv_score[meta$has_cnv_score > cnv_threshold] <- "CNV"

  # we know that some immune cells might have false positive CNV 
  meta$has_cnv_score[meta$has_cnv_score %in% c(0:cnv_threshold)] <- "no CNV"
  infercnv_obj <- AddMetaData(infercnv_obj, meta$has_cnv_score, col.name =  "has_cnv_score") 
  p[["infercnv_obj"]] <- infercnv_obj

@@

return(p)

}

```

@@

```{r}

result_dir <- file.path(module_base, "results", params$sample_id, "06_infercnv")

```

@@

```{r}

srat <- readRDS(file.path(module_base, "results", params$sample_id, glue::glue("02b-fetal_kidney_label-transfer_", params$sample_id, ".Rds")))

```

@@

```{r fig.width=20, fig.height=5, out.width='100%', out.height='100%'}
infercnv_obj <- list()
for(reference_value in c("reference-none", "reference-immune", "reference-endothelium", "reference-both")){
  infercnv_obj[[reference_value]] <- readRDS(file.path(result_dir, glue::glue(reference_value, "_HMM-no") , glue::glue("06_infercnv_", params$sample_id, "_", reference_value, "_HMM-no.rds")))
  print(Do_CNV_heatmap(object = srat, infercnv_obj = infercnv_obj[[reference_value]], group.by = "fetal_kidney_predicted.compartment", reference_value = reference_value))
}

```
These heatmaps emphasize the importance of the selection of normal cells prior the inference of CNV.
The normal reference should contain as much cell types as possible, in order to minimize false positive CNV.
In our case, we should take immune and entodethial cells when possible.

@@

```{r fig.width=20, fig.height=5, out.width='100%', out.height='100%', warning=FALSE}

for(reference_value in c("reference-none", "reference-immune", "reference-endothelium", "reference-both")){
  srat <- Do_CNV_score(srat,infercnv_obj = infercnv_obj[[reference_value]], reference_value)
  p1 <- visualize_feature(srat, feature =  glue::glue("CNV-score_", reference_value) , group.by = "fetal_kidney_predicted.compartment" )
  p2 <- visualize_density(srat, features=glue::glue("CNV-score_", reference_value), group.by = "fetal_kidney_predicted.compartment")
  print(p1 + p2 + plot_layout(ncol = 3, widths = c(1,2,2)))
}

```

@@

```{r}

#### Load the `Seurat` object
srat_i3 <- readRDS(file.path(module_base, "results", params$sample_id, glue::glue("06_infercnv_HMM-i3_", params$sample_id, "_reference-both.rds")))

```

@@

```{r}

srat_i6 <- readRDS(file.path(module_base, "results", params$sample_id, glue::glue("06_infercnv_HMM-i6_", params$sample_id, "_reference-both.rds")))

```

@@

meta <- seurat@meta.data[colnames(srat),]

# merge the new info into the srat metadata
srat_pull <- AddMetaData(srat, meta[,grepl("_cnv_chr", colnames(meta))], col.name = colnames(meta[grepl("_cnv_chr", colnames(meta))]))
```

@@

```{r fig.width=20, fig.height=8, out.width='100%'}

p1 <- SCpubr::do_DimPlot(pull$infercnv_obj, "has_cnv_score", plot.title = "inter-patient reference, HMM-i3", reduction = "umap")

@@

p1 + p2 + p3 + plot_layout(ncol = 3)

```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warnings = FALSE
)
```

@@

```{r}
subdiagnosis <- readr::read_tsv(
  file.path("..", "..", "..", "data", "current", params$scpca_project_id, "single_cell_metadata.tsv"),
  show_col_types = FALSE
) |>
  dplyr::filter(scpca_sample_id == params$sample_id) |>
  dplyr::pull(subdiagnosis)
```

@@

```{r}
library(infercnv)
library(SCpubr)
library(ggplot2)
library(patchwork)
library(Seurat)
```

@@

```{r}
Do_CNV_heatmap <- function(object, infercnv_obj, group.by, reference_value) {
  out <- SCpubr::do_CopyNumberVariantPlot(
    sample = object,
    infercnv_object = infercnv_obj,
    using_metacells = FALSE,
    chromosome_locations = SCpubr::human_chr_locations,
    return_object = FALSE,
    group.by = group.by
  )
  out <- out +
    ggtitle(glue::glue("Copy Number Variant Plot, ", reference_value)) +
    ylab(label = "")

  return(out)
}
```

@@

```{r}
Do_CNV_score <- function(seurat_oject, infercnv_obj, reference_value) {
  score <- apply(infercnv_obj@expr.data, 2, function(x) {
    sum(x < 0.95 | x > 1.05) / length(x)
  })

@@

```{r }
visualize_metadata <- function(object, meta, group.by) {
  if (is.numeric(object@meta.data[, meta])) {
    d <- SCpubr::do_FeaturePlot(object,
      features = meta,
      pt.size = 0.2,
      legend.width = 0.5,
      legend.length = 5,
      legend.position = "right"
    ) + ggtitle(meta)
    b <- SCpubr::do_ViolinPlot(object,
      features = meta,
      ncol = 1,
      group.by = group.by,
      legend.position = "none"
    )

    return(d + b + plot_layout(ncol = 2, widths = c(2, 4)))
  } else {
    d <- SCpubr::do_DimPlot(object, reduction = "umap", group.by = group.by, label = TRUE, repel = TRUE) + ggtitle(paste0(meta, " - umap")) + theme(text = element_text(size = 18))
    b <- SCpubr::do_BarPlot(
      sample = object,
      group.by = meta,
      split.by = group.by,
      position = "fill",
      font.size = 10,
      legend.ncol = 3
    ) +
      ggtitle("% cells") +
      xlab(print(group.by)) +
      theme(text = element_text(size = 18))
    return(d + b + plot_layout(ncol = 2, widths = c(2, 4)))
  }
}
```

@@

```{r }
visualize_feature <- function(object, feature, group.by) {
  d <- SCpubr::do_FeaturePlot(object,
    features = feature,
    pt.size = 0.2,
    legend.width = 0.5,
    legend.length = 5,
    legend.position = "right"
  ) + ggtitle(as.character(feature))
  b <- SCpubr::do_ViolinPlot(object,
    features = feature,
    ncol = 1,
    group.by = group.by,
    legend.position = "none",
    assay = "SCT"
  ) + ylab(as.character(feature))

  return(d + b + plot_layout(ncol = 2, widths = c(2, 4)))
}
```

@@

```{r }
visualize_density <- function(object, features, group.by) {
  d <- SCpubr::do_RidgePlot(object,
    feature = features,
    group.by = group.by,
    legend.position = "none",
    assay = "SCT"
  ) + ylab(features)

  return(d)
}
```

@@

```{r}
wrapper_explore_hmm <- function(infercnv_obj, cnv_threshold = 1) {
  p <- list()

@@

  #### Feature plot and repartition of the CNV per chromosome
  for (i in 1:22) {
    tmp1 <- visualize_feature(infercnv_obj, feature = glue::glue("proportion_cnv_chr", i), group.by = "fetal_kidney_predicted.compartment")
    tmp2 <- visualize_density(infercnv_obj, features = glue::glue("proportion_cnv_chr", i), group.by = "fetal_kidney_predicted.compartment")
    p[[glue::glue("proportion_cnv_chr", i)]] <- tmp1 + tmp2 + plot_layout(ncol = 3, widths = c(1, 2, 2))
  }

  #### DotPlot
  p[["Dotplot"]] <- DotPlot(infercnv_obj, features = colnames(infercnv_obj@meta.data)[grepl("has_cnv_chr", colnames(infercnv_obj@meta.data))], group.by = "fetal_kidney_predicted.compartment", assay = "SCT") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

  #### CNV score
  meta <- infercnv_obj@meta.data |>
    mutate(has_cnv_score = rowSums(meta[, grepl("has_cnv_chr", colnames(meta))])) |>
    mutate(has_cnv_score = case_when(
      has_cnv_score > cnv_threshold ~ "CNV",
      has_cnv_score <= cnv_threshold ~ "no CNV"
    ))

  infercnv_obj <- AddMetaData(infercnv_obj, meta$has_cnv_score, col.name = "has_cnv_score")
  p[["infercnv_obj"]] <- infercnv_obj

@@

  return(p)
}
```

@@

```{r}
result_dir <- file.path(module_base, "results", params$sample_id, "06_infercnv")
```

@@

```{r}
srat <- readRDS(file.path(module_base, "results", params$sample_id, glue::glue("02b-fetal_kidney_label-transfer_", params$sample_id, ".Rds")))
```

@@

```{r fig.width=20, fig.height=5, out.width='100%', out.height='100%'}
infercnv_obj <- list()
for (reference_value in c("reference-none", "reference-immune", "reference-endothelium", "reference-both")) {
  infercnv_obj[[reference_value]] <- readRDS(file.path(result_dir, glue::glue(reference_value, "_HMM-no"), glue::glue("06_infercnv_", params$sample_id, "_", reference_value, "_HMM-no.rds")))
  print(Do_CNV_heatmap(object = srat, infercnv_obj = infercnv_obj[[reference_value]], group.by = "fetal_kidney_predicted.compartment", reference_value = reference_value))
}
```
These heatmaps emphasize the importance of the selection of normal cells prior the inference of CNV.
The normal reference should contain as much cell types as possible, in order to minimize false positive CNV.
In our case, we should take immune and entodethial cells when possible.

@@

```{r fig.width=20, fig.height=5, out.width='100%', out.height='100%', warning=FALSE}
for (reference_value in c("reference-none", "reference-immune", "reference-endothelium", "reference-both")) {
  srat <- Do_CNV_score(srat, infercnv_obj = infercnv_obj[[reference_value]], reference_value)
  p1 <- visualize_feature(srat, feature = glue::glue("CNV-score_", reference_value), group.by = "fetal_kidney_predicted.compartment")
  p2 <- visualize_density(srat, features = glue::glue("CNV-score_", reference_value), group.by = "fetal_kidney_predicted.compartment")
  print(p1 + p2 + plot_layout(ncol = 3, widths = c(1, 2, 2)))
}
```

@@

```{r}
#### Load the `Seurat` object
srat_i3 <- readRDS(file.path(module_base, "results", params$sample_id, glue::glue("06_infercnv_HMM-i3_", params$sample_id, "_reference-both.rds")))
```

@@

```{r}
srat_i6 <- readRDS(file.path(module_base, "results", params$sample_id, glue::glue("06_infercnv_HMM-i6_", params$sample_id, "_reference-both.rds")))
```

@@

meta <- seurat@meta.data[colnames(srat), ]

# merge the new info into the srat metadata
srat_pull <- AddMetaData(srat, meta[, grepl("_cnv_chr", colnames(meta))], col.name = colnames(meta[grepl("_cnv_chr", colnames(meta))]))
```

@@

```{r fig.width=20, fig.height=8, out.width='100%'}
p1 <- SCpubr::do_DimPlot(pull$infercnv_obj, "has_cnv_score", plot.title = "inter-patient reference, HMM-i3", reduction = "umap")

@@

p1 + p2 + p3 + plot_layout(ncol = 3)
```
